---
title: "Portfolio1_sleepstudy"
author: "Kristine"
date: "3/2/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("tidyverse")
#import data
sleepstudy <- read.csv("sleepstudy.csv")

df337 <- filter(sleepstudy, Subject == 337)

summary(lm(Reaction~Days, data = df337))


```
#Question 1.a Report the F.statistics
```{r}
#1.a 112.1
```

#Question 1.b How many degrees of freedom does the relevant F-distribution have?
```{r}
#1.b df(1,8)
```

#Question 1.c At which F-value does a regression with the distribution become statistically significant (p<0.05)?
```{r}
#Getting the F-score for the 95% significance cutoff using the quantile function qf()


cutoff<-qf(0.95,1,8)
'p=0.05 cutoff value with df(2,12)'
cutoff


#1.c F=5.3

```

#Question 1.d Make a plot of the F-distribution 
```{r}
#Another way of doing it 
pacman::p_load(sjPlot)
library(sjPlot)
dist_f(f=0, deg.f1 =1, deg.f2 = 8, xmax = 10)

```


##Task 2 
For all participant in the experiment:
2.a Find the coeffcients (slope and intercept) for the regression for reaction time as a function of days of sleep deprivation (a hint for the solution: use group_by() in tidyverse or this function here: https://stat.ethz.ch/R-manual/R-devel/library/nlme/html/lmList.html, hint2: pool=FALSE

```{r}
library(nlme)
fm1 <- nlme::lmList(Reaction ~ Days | Subject, data = sleepstudy, pool=FALSE)
summary(fm1)
coef(fm1)
```


2.b: Combine both scatter plot and regression line in the same ﬁgure. You may also include all participants in one plot. 

```{r}

ggplot(sleepstudy, aes(y=Reaction, x=Days, color = as.character(Subject)))+
  geom_point() + 
  geom_smooth(method = lm, se=FALSE) + 
  ggtitle("Reaction time as a function of days of sleep deprivation")


```


2.c: Collect and report the inferential statistics for each participant in a table using t-statistics, including t-value, df and p-value.

```{r}

#we make a data frame containing the coeffecients from our models
coef <- coef(summary(fm1))
fm1_stat <- as_tibble(coef)

#We rename the column names 
names(fm1_stat)[names(fm1_stat) == "Estimate.Days"] <- "Slopes"
names(fm1_stat)[names(fm1_stat) == "Estimate.(Intercept)"] <- "Intercept"
names(fm1_stat)[names(fm1_stat) == "Pr(>|t|).Days"] <- "P-value slopes"

view(fm1_stat)

```


2.d: How many individual participants display a statistically signiﬁcant eﬀect of sleep deprivation (p-values uncorrected for mulitple comparisons)? 


##Task 3
Across participants: 
3.a: Use the slopes you found for each participant in exercise 2 as a new dataset. Test the hypothesis that the slopes are larger than zero against the null-hypothesis that the slopes are zero (i.e. no differences in response time exist as a function of time).

```{r}
#Create a dataframe with the slopes
#df_slope <- coef(fm1)
#df_slope <- rename(df_slope, c("Days"="Slope", "(Intercept)"="Intercept"))

#Compare the slopes to the null hypothesis that the slopes equal zero, meaning that there is no relation between days of sleep deprivation and rection time
#t.test(df_slope$Slope, mu = 0.0)
```

Assumptions - the data should be normally distributed 
- Manon checked that using a histogram and a Q-Q plot. 

3.b: Justify your use of test statistics.

3.c: Report inferential statistics.

3.d: Make a plot with the mean reaction time and standard errror bars for each day across participants and plot the averaged regression line in the same ﬁgure

```{r}
#ggplot(sleepstudy, aes(x=Days,y=Reaction))+
  #geom_smooth(method=lm)+
  #geom_errorbar(stat="summary", fun.data=mean_se, width=0.3, color ="#000000")+
  #stat_summary()
```


