---
title: "Portfolio1_sleepstudy"
author: "Kristine"
date: "3/2/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(pacman)
install.packages("tidyr")
library(tidyverse)
library(ggplot2)
#import data
sleepstudy <- read.csv("sleepstudy.csv")

df337 <- filter(sleepstudy, Subject == 337)

summary(lm(Reaction~Days, data = df337))


```
#Question 1.a Report the F.statistics
```{r}
#1.a F(1,8)=112.1
```

#Question 1.b How many degrees of freedom does the relevant F-distribution have?
```{r}
#1.b df(1,8)
```

#Question 1.c At which F-value does a regression with the distribution become statistically significant (p<0.05)?
```{r}
#Getting the F-score for the 95% significance cutoff using the quantile function qf()


cutoff<-qf(0.95,1,8)
'p=0.05 cutoff value with df(2,12)'
cutoff


#1.c F=5.3

```

#Question 1.d Make a plot of the F-distribution 
```{r}
nn <- seq(0,120, len = 1000)
fdist <- df(nn,1,8)
plot(nn,fdist, type='l',xlab='F value - df(1,8)',ylab='Density')

#Another way of doing it 
install.packages("tidyr")
library(sjPlot)
dist_f(f=0, deg.f1 =1, deg.f2 = 8, xmax = 10)

```


##Task 2 
For all participant in the experiment:
#2.a Find the coeffcients (slope and intercept) for the regression for reaction time as a function of days of sleep deprivation (a hint for the solution: use group_by() in tidyverse or this function here: https://stat.ethz.ch/R-manual/R-devel/library/nlme/html/lmList.html, hint2: pool=FALSE

We create a model with reaction time as dependent variable and days of sleep deprivation as the independent variable. A model is made for each subject in the study.

```{r}
fm1 <- nlme::lmList(Reaction ~ Days | Subject, sleepstudy, pool = FALSE)
summary(fm1)
coef(fm1)
```
The coeffecients for each model is displayed in the table.

#2.b: Combine both scatter plot and regression line in the same ﬁgure. You may also include all participants in one plot.

We make a plot showing the regression line of each model

```{r}
ggplot(sleepstudy, aes(y=Reaction, x=Days, color = as.character(Subject)))+
  geom_point() + 
  geom_smooth(method = lm, se=FALSE) + #fit the best straight line to the data
  ggtitle("Reaction time as a function of days of sleep deprivation")

```


#2.c: Collect and report the inferential statistics for each participant in a table using t-statistics, including t-value, df and p-value.

```{r}

#we create a list of the models. 

fm2 <- nlme::lmList(Reaction ~ Days | Subject, sleepstudy, pool = FALSE)

#we make a data frame containing the coeffecients from our models
df_coef <- coef(summary(fm1))
df_stat <- as_tibble(df_coef)

#We rename the column names 
names(df_stat)[names(df_stat) == "Estimate.Days"] <- "Slopes"
names(df_stat)[names(df_stat) == "Estimate.(Intercept)"] <- "Intercept"
names(df_stat)[names(df_stat) == "Pr(>|t|).Days"] <- "p-value slopes"

view(df_stat)
  
```


#2.d: How many individual participants display a statistically signiﬁcant effect of sleep deprivation (p-values uncorrected for mulitple comparisons)? 
We use the select function to look at the relevant p-values. Furthermore we use the filter function to show the p-values that are statistically signinficant (p<0.05)

```{r}
p_signif <- select(df_stat, "p-value slopes")

p_signif %>% filter("p-value slopes"<0.5)

view(p_signif)
```

As we cannot make our code above work, we have created individual models for each partcipant. The p-values are taken from the summary output of the models. 
- furthermore the insignificant p-values can be found in the table displayed in 2.c, by looking in the "p-value slopes" column and  number 4, 5, 6, 9. 

```{r}

df308 <- filter(sleepstudy, Subject == 308)
summary(lm(Reaction~Days, data = df308))
#p-value = 0.0033

df309 <- filter(sleepstudy, Subject == 309)
summary(lm(Reaction~Days, data = df309))
#p-value = 0.049

df310 <- filter(sleepstudy, Subject == 310)
summary(lm(Reaction~Days, data = df310))
#p-value = 0.002

df330 <- filter(sleepstudy, Subject == 330)
summary(lm(Reaction~Days, data = df330))
#p-value = 0.255

df331 <- filter(sleepstudy, Subject == 331)
summary(lm(Reaction~Days, data = df331))
#p-value = 0.076

df332 <- filter(sleepstudy, Subject == 332)
summary(lm(Reaction~Days, data = df332))
#p-value = 0.191

df333 <- filter(sleepstudy, Subject == 333)
summary(lm(Reaction~Days, data = df333))
#p-value = 0.0002

df334 <- filter(sleepstudy, Subject == 334)
summary(lm(Reaction~Days, data = df334))
#p-value = 0.0006

df335 <- filter(sleepstudy, Subject == 335)
summary(lm(Reaction~Days, data = df335))
#p-value = 0.051

df337 <- filter(sleepstudy, Subject == 337)
summary(lm(Reaction~Days, data = df337))
#p-value = 0.000005

df349 <- filter(sleepstudy, Subject == 349)
summary(lm(Reaction~Days, data = df349))
#p-value = 0.000023

df350 <- filter(sleepstudy, Subject == 350)
summary(lm(Reaction~Days, data = df350))
#p-value = 0.000086

df351 <- filter(sleepstudy, Subject == 351)
summary(lm(Reaction~Days, data = df351))
#p-value = 0.033

df352 <- filter(sleepstudy, Subject == 352)
summary(lm(Reaction~Days, data = df352))
#p-value = 0.0013

df369 <- filter(sleepstudy, Subject == 369)
summary(lm(Reaction~Days, data = df369))
#p-value = 0.00019

df370 <- filter(sleepstudy, Subject == 370)
summary(lm(Reaction~Days, data = df370))
#p-value = 0.00014

df371 <- filter(sleepstudy, Subject == 371)
summary(lm(Reaction~Days, data = df371))
#p-value = 0.010

df372 <- filter(sleepstudy, Subject == 372)
summary(lm(Reaction~Days, data = df372))
#p-value = 0.000017
```

We found that 14 out of 18 participants showed a statistically signiﬁcant effect of sleep deprivation. Subject 330 (p=0.255), 331 (p=0.076), 332 (p=0.191), 335 (p=0.051) had unsignificant p-values. 

##Task 3
Across participants: 
3.a: Use the slopes you found for each participant in exercise 2 as a new dataset. Test the hypothesis that the slopes are larger than zero against the null-hypothesis that the slopes are zero (i.e. no differences in response time exist as a function of time).

```{r}
#Create a dataframe with the slopes
df_slope <- coef(fm1)
df_slope <- rename(df_slope, c("Days"="Slope", "(Intercept)"="Intercept"))

#Compare the slopes to the null hypothesis that the slopes equal zero, meaning that there is no relation between days of sleep deprivation and rection time
t.test(df_slope$Slope, mu = 0.0)
```

Assumptions - the data should be normally distributed 
- Manon checked that using a histogram and a Q-Q plot. 

3.b: Justify your use of test statistics.

3.c: Report inferential statistics.

3.d: Make a plot with the mean reaction time and standard errror bars for each day across participants and plot the averaged regression line in the same ﬁgure

```{r}
ggplot(sleepstudy, aes(x=Days,y=Reaction))+
  geom_smooth(method=lm)+
  geom_errorbar(stat="summary", fun.data=mean_se, width=0.3, color ="#000000")+
  stat_summary()
```


